The interactions of rational, pragmatic agents  \n lead to efficient language structure and use

### 2019-02-02 paper/paper.Rmd
more tweaks
MF last changes and spell-checking.

Zipf articulated what is now considered a *functionalist* approach to language science -- analyzing language structure and use in terms of efficiency. Such an approach might reframe our opening question as follows: how does having property \textit{x} make using language $\ell$ more or less useful for communication? Among others, this efficiency-based framing has produced a rich set of theoretical and empirical targets exploring semantic typology [@RegierKempKay2015a], properties such as ambiguity [@Piantadosi2011a] and compositionality [@KirbyGriffithsSmith2014a], and the efficient use of reduction and redundancy in production [@LevyJaeger2007a; @GenzelCharniak2002a].\par

The approaches above typically posit efficiency measures motivated by information-theoretic principles, but not directly grounded in language use by interacting agents. In this work, we derive a measure of efficiency from first principles of \textit{efficient language use} and show how this can lead to communicatively efficient systems. We also demonstrate how assumptions about interlocutors impacts how language properties are used efficiently. In this way, we bridge questions of design and use in a single framework.\par

Functionalist theories commonly frame language efficiency in terms of a fundamental effort-assymetry underlying everyday communication: what is “hard” for a speaker is likely different than what is “hard” for a listener. Zipf described this as follows -- purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of a single, low-cost word. Given such a language, the full set of potential meanings would be conveyed using only that word, i.e. $\ell_{speaker}^*$ would be fully ambiguous and all possible meanings would need to be disambiguated by a listener. From the standpoint of listener effort, an optimal language $\ell_{listener}^*$ would map all possible meanings to distinct words, removing a listener's need to disambiguate. In this example, speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*. Clearly, natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$. Zipf proposed that the particular lexicon-level properties he observed were a result of these competing forces -- the pressure to jointly minimize speaker and listener effort.\par

@@

Adopting reference games as a basic unit of analysis suggests that optimization may take place at the level of conversation. Importantly, Zipf's conception of speaker and listener effort may be connected to how language is used; in particular, whether interlocutors engage in pragmatic reasoning during conversation. Under a Gricean treatment of pragmatics, speakers and listeners follow a set of conversational maxims in which they cooperate to transfer information [@Grice1975a]. These maxims appear to emerge from efficiency concerns, however (Horn, 1984). We formalize this connection -- showing how system-level efficiencies can emerge from local interaction behavior of pragmatic agents. Our claim is that to understand an "efficient" property of a system it is essential that we consider how that property is *used* efficiently.\par

We provide a case study for this approach, in which functionalist regularities emerge from the dynamics of pragmatic communication. We choose a property of languages that could, in principle, vary freely, but shows strong regularities across languages. The explanadum is why this regularity holds. We examine ambiguity as our property, extending ideas by Piantadosi et al. [-@Piantadosi2011a]. We define a novel measure of efficiency that depends on the interactional behavior of speaker and listener agents. We adopt the reference game as our primary unit of interaction and model language users with the Rational Speech Act (RSA) framework -- a computional model of language use, which is nicely attested by experimental data on interaction. Using these ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency (Simulation 1). Further, we show how ambiguity is used efficiently during local, in-the-moment interactions (Simulation 2). Put differently, these simulations examine efficiency from two angles -- in the first we vary languages, fixing agents, and search for efficient designs. In the second we vary agents, fixing language, and examine efficient use. \par

@@

### Reference games
Zipf’s example of optimal speaker- and listener-languages took the form of a reference game. We adopt that formulation here, assuming these communication games as our basic unit of analysis. In this framework, speakers and listeners are aware of a set of objects $M$ (*meanings*) and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1). Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$. Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings are analogous to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}. We consider a set of contexts $C$ with an associated prior $P(C)$. Each context $c\in C$ describes a different distribution over meanings e.g. $p(M|C=c_i) \neq p(M|C=c_j)$. Finally, we consider a set of communicative events $e \in E$ where $<u, m, c> = e$ is an utterance-meaning-context triple.\par

### Languages
A language $\ell$ defines the set of semantic mappings between utterances and meanings. For example, Figure 1 contains four utterances $U = \{\text{"blue"}, \text{"green"}, \text{"square"}, \text{"circle"}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}$. The boolean matrix describes the literal semantics of the language. We define a language as "ambiguous" if there is some utterance $u \in U$ which can apply to multiple meanings (i.e. $|[[u_i]]| > 1$)\footnote{We use double brackets $[[\dots]]$ to represent denotation.}. In Figure 1 both the words "square" and "green" are ambigous so we would say that $\ell$ contains ambiguity.

### Speakers and listeners
The Rational Speech Act framework (RSA) is a computational-level theory of pragmatic language use, which has produced good fit to human communication behavior across a range of language phenomena [@FrankGoodman2012a; @GoodmanFrank2016a]. RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a]. We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI} and @GoodmanFrank2016a for a review).\par

@@

We assume that speaker effort is related to the surprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterances costs as independent from context (ie. $p(u|c)p(c)=p(u)p(c)$).} -- intuitively, the number of bits needed to encode the utterance $u$. This particular formalization of speaker-cost is general enough to accommodate a range of cost instantiations, such as production difficulty via articulation effort, cognitive effort related to lexical access, or others [@BennettGoodman2015a].\par

@@

Importantly, we assume that events $e = <u, m, c>$ are sampled according the to following generative model -- some context occurs in the world with probability $P(C=c)$. Within this context, an object $m$ occurs with probability $p(m|c)$. The speaker attempts to refer to that object by sampling from her conditional distribution $S(u|m, c; \ell)$ (i.e. $e \sim p(c)p(m|c)S(u|m, c; \ell)$). From these ingredients it is possible to derive the following objective between the speaker and listener distributions (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.1} for complete derivation).
\begin{equation}
\begin{split}
  = \mathbb{E}_{c \sim P(C)}[H_{cross}(P_{speaker}, P_{listener} | c; \ell)]\\
\end{split}
\end{equation}
From an information-theoretic perspective this objective is intuitive: $H_{cross}$ denotes the Cross-Entropy (CE), a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs. In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$\footnote{Note that in the single context case $|C|=1$ this objective is simply the speaker-listener Cross-Entropy.}. In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listeners think.

@@

```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2.8, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity. Horizontal axis shows the context-size (1-4) in each condition. Optimal language under CE objective (red). Speaker-optimal (blue). Listener-optimal (green). Error bars represent 95 percent confidence intervals. Panel (B), example CE-optimal language (ambiguous) under in a four-condition context. Panel (C), example CE-optimal language (unambiguous) in a single-condition context."}
img <- png::readPNG("figs/fig1.png")
grid::grid.raster(img)
```

@@

For comparison, we also examine properties of optimal languages under two additional objectives. Zipf [-@Zipf1949a] proposed that the speaker-optimal language $\ell_{speaker}^*$ would minimize speaker effort and the listener-optimal language $\ell_{listener}^*$ would minimize listener effort. We define these objectives using the first and second half of equation 1 (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 2.2.}).\par

@@

Imagine a scenario in which a reader is beginning a news article. While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved. In this setting, using a low-cost, but ambiguous referring expression (say a pronoun like "he") early may lead to misunderstanding if context is not informative. But, if by a later position enough contextual information has accumulated, it may be efficient to use the ambiguous expression. We pursue this general framework in Simulation 2 -- examining when in a discourse using ambigity is efficient. Note that we will consider “context” as analogous to a “topic” of conversation. \par

@@

We conduct $N=600$ simulations, generating discourses of length $|D|=30$ utterances, comparing three speaker models ($n=200$ each). We consider a single language $\ell$\footnote{See SI for the matrix notation of this langauge.} with $|U|=6$ and $|M|=4$ in which two of the utterances are ambiguous and lower cost than the unambiguous utterances. (Note that use of this particular language is not essential -- the results are broadly generalizable to languages that contain ambiguity, but exploring this space is computationally expensive.)\par

## Speaker agents
We vary the degree to which agents can use context for disambiguation. We consider three types of speaker models. Our \textit{Full pragmatics} agent, models a speaker who reasons about her listener and also has complete recall of the set of utterances in the discourse $D$. This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discourse, but can infer it over the discourse. We compare two baseline models. The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to the discourse history. The second, a \textit{No pragmatics} baseline speaker does not consider a listener at all, but produces utterances according to the underlying langauge semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see \href{https://github.com/benpeloquin7/zipf_principles/blob/master/paper/supplementary_materials.pdf}{SI 3}).\par

@@

Figure 3, panel (A) shows the empirical probability a speaker uses an ambiguous utterance as a function of discourse position. The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse. The \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely. But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds. Panel (C) tracks our CE objective for each model over the discourse. Note that the objective decreases for all three models, primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$). However, the objective declines more quickly under the \textit{Full} and \textit{Partial pragmatic} speakers as listener agents are better able to infer the correct context. Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs. Panel (B) tracks speaker effort, which remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines. But, effort declines in the \textit{Full pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par

@@

How do the competing pressures of speakers and listeners give rise to the distributional regularities found in natural language? Zipf [-@Zipf1949a] proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon. We explored the interactions of rational pragmatic agents as a framework for understanding efficient language structure and use. We focused on an argument on the communicative function of ambiguity [@Piantadosi2011a], deriving a novel speaker-listener Cross-Entropy objective for measuring the efficiency of linguistic systems from first principles of efficient language use. In Simulation 1 we showed that optimal languages are more likely to contain ambiguous material when context is informative. In Simulation 2 we showed how rational pragmatic agents use ambiguous material efficiently in conversation.\par

A limitation of the current work is an analysis of exactly how the CE objective compares to existing measures. For example, previous work has described competing speaker-listener pressures in terms of a trade-off of simplicity and informativeness [@KempRegier2012a] or expressivity and compressibility [@SmithTamarizKirby2013a] to explain linguistic regularities. Future work should assess the degree to which we can derive the same properties as previous studies using our current framework. In line with this idea, a primary focus of the current work is to provide a case-study on the productivity of assessing functionalist theories in repeated reference games between rational pragmatic agents. We hope this framework can serve as a domain general tool to assess the range of functionalist theories examining efficient language-structure and use.\par

\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI: \href{https://bit.ly/2DOnjI4}{\url{https://bit.ly/2DOnjI4}}.}}} \vspace{1em}
\noindent

>>>>

Zipf articulated what is now considered a *functionalist* approach to language science -- analyzing language structure and use in terms of efficiency. Such an approach might re-frame our opening question as follows: how does having property \textit{x} make using language $\ell$ more or less useful for communication? This efficiency-based framing has produced a rich set of theoretical and empirical targets exploring semantic typology [@RegierKempKay2015a], properties such as ambiguity [@Piantadosi2011a] and compositionality [@KirbyGriffithsSmith2014a], and the efficient use of reduction and redundancy in production [@LevyJaeger2007a; @GenzelCharniak2002a].\par

The approaches above typically posit efficiency measures that are motivated by information-theoretic principles, but they typically do not ground out in language use by interacting agents. In this work, we derive a novel objective function from first principles of \textit{rational language use} and show how optimizing this objective can lead to communicatively efficient systems. We also demonstrate that assumptions about interlocutors impact whether language properties are used efficiently. In this way, we integrate questions of language design and language use in a single framework.\par

Functionalist theories commonly frame language efficiency in terms of a fundamental effort-asymmetry underlying everyday communication: what is “hard” for a speaker is likely different than what is “hard” for a listener. Zipf described this as follows: purely from the standpoint of speaker effort, an optimal language $\ell_{speaker}^*$ would tend toward a vocabulary of a single, low-cost word. Given such a language, the full set of potential meanings would be conveyed using only that word, i.e. $\ell_{speaker}^*$ would be fully ambiguous and all possible meanings would need to be disambiguated by a listener. From the standpoint of listener effort, an optimal language $\ell_{listener}^*$ would map all possible meanings to distinct words, removing a listener's need to disambiguate. In this example, speaker effort is related to *production cost* and listener effort to *understanding or disambiguation cost*. Clearly, natural languages fall between the two extremes of $\ell_{speaker}^*$ and $\ell_{listener}^*$. Zipf proposed that the particular lexicon-level properties he observed were a result of optimization based on these competing forces -- the pressure to jointly minimize speaker and listener effort.\par

@@

Adopting reference games as a basic unit of analysis suggests that optimization may take place at the level of conversation. Importantly, Zipf's conception of speaker and listener effort should be connected to how language is used; in particular, whether interlocutors engage in pragmatic reasoning during conversation. Under a Gricean treatment of pragmatics, speakers and listeners follow a set of conversational maxims in which they cooperate to transfer information [@Grice1975a]. These maxims appear to emerge from efficiency concerns, however (Horn, 1984). We formalize this connection -- showing how system-level efficiencies can emerge from local interaction behavior of pragmatic agents. Our claim is that to understand an "efficient" property of a system it is essential that we consider how that property is *used* efficiently.\par

We provide a case study for this approach, in which functionalist regularities emerge from the dynamics of pragmatic communication. We choose a property of languages that could, in principle, vary freely, but shows strong regularities across languages. The explanandum is why this regularity holds. We examine ambiguity as our property, extending ideas by Piantadosi et al. [-@Piantadosi2011a]. We define a novel measure of efficiency that depends on the interactional behavior of speaker and listener agents. We adopt the reference game as our primary unit of interaction and model language users with the Rational Speech Act (RSA) framework -- a computational model of language use, which is supported by experimental data on interaction. Using these ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency (Simulation 1). Further, we show how ambiguity is used efficiently during local, in-the-moment interactions (Simulation 2). Put differently, these simulations examine efficiency from two angles -- in the first we vary languages, fixing agents, and search for efficient language designs. In the second we vary agents, fixing language, and examine efficient use. \par

@@

### Reference games
Zipf’s example of optimal speaker- and listener-languages took the form of a reference game. We adopt that formulation here, assuming these communication games as our basic unit of analysis. In this framework, speakers and listeners are aware of a set of objects $M$ (*meanings*) and are knowledgeable about the set of possible signals $U$ (*utterances*) that can be used to refer to a given meaning (see Figure 1). Utterances may have different relative costs, operationalized via a prior over utterances $P(U)$. Similarly, meanings differ in the relative degree to which they need to be talked about, operationalized as a prior over meanings $P(M)$\footnote{The prior over meanings is equivalent to the \textit{need probabilities} assumed in previous work (Regier, Kemp \& Kay (2015).}. We consider a set of contexts $C$ with an associated prior $P(C)$. Each context $c\in C$ describes a different distribution over meanings e.g. $p(M|C=c_i) \neq p(M|C=c_j)$. Finally, we consider a set of communicative events $e \in E$ where $<u, m, c> = e$ is an utterance-meaning-context triple.\par

### Languages
A language $\ell$ defines the set of semantic mappings between utterances and meanings. For example, Figure 1 contains four utterances $U = \{\text{"blue"}, \text{"green"}, \text{"square"}, \text{"circle"}\} \text{ and three meanings }M = \{\text{green-square}, \text{blue-square}, \text{green-circle}\}$. The boolean matrix describes the literal semantics of the language. We define a language as "ambiguous" if there is some utterance $u \in U$ which can apply to multiple meanings (i.e. $|[[u_i]]| > 1$)\footnote{We use double brackets $[[\dots]]$ to represent denotation.}. In Figure 1 both the words "square" and "green" are ambiguous so we would say that $\ell$ contains ambiguity.

### Speakers and listeners
The Rational Speech Act framework (RSA) is a computational-level theory of pragmatic language use, which has produced good fit to human communication behavior across a range of language phenomena [@FrankGoodman2012a; @GoodmanFrank2016a]. RSA is a formalization of essential Gricean pragmatic principles -- agents reason about one another and their shared context [@Grice1975a]. We adopt RSA as our representational framework to model Gricean (rational and pragmatic) speaker-listeners in the reference game setting (see \href{https://bit.ly/2RBSGcU}{SI}).\par

@@

We assume that speaker effort is related to the surprisal of an utterance in a particular context\footnote{In the current set of simulations we consider utterances costs as independent from context (i.e.. $p(u|c)p(c)=p(u)p(c)$).} -- intuitively, the number of bits needed to encode the utterance $u$. This particular formalization of speaker-cost is general enough to accommodate a range of cost instantiations, such as production difficulty via articulation effort, cognitive effort related to lexical access, or others [@BennettGoodman2015a].\par

@@

Importantly, we assume that events $e = <u, m, c>$ are sampled according the to following generative model -- some context occurs in the world with probability $P(C=c)$. Within this context, an object $m$ occurs with probability $p(m|c)$. The speaker attempts to refer to that object by sampling from her conditional distribution $S(u|m, c; \ell)$ (i.e. $e \sim p(c)p(m|c)S(u|m, c; \ell)$). From these ingredients it is possible to derive the following objective between the speaker and listener distributions (see \href{https://bit.ly/2RBSGcU}{SI 2.1} for complete derivation).
\begin{equation}
\begin{split}
  = \mathbb{E}_{c \sim P(C)}[H_{cross}(P_{speaker}, P_{listener} | c; \ell)]\\
\end{split}
\end{equation}
From an information-theoretic perspective this objective is intuitive: $H_{cross}$ denotes the Cross-Entropy (CE), a measure of dissimilarity between two distributions -- the average number of bits required to communicate under one distribution, given that the “true” distribution differs. In our case, we have an expectation over this term -- the expected difference between the distributions assumed by the speaker $P_{speaker}$ and listener $P_{listener}$ given a set of contexts $C$\footnote{Note that in the single context case $|C|=1$ this objective is simply the speaker-listener Cross-Entropy.}. In other words, an "efficient" language $\ell$ minimizes the distance between what speakers and listeners think.

@@

```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2.8, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity. Horizontal axis shows the context-size (1-4) in each condition. Optimal language under CE objective (red). Speaker-optimal (blue). Listener-optimal (green). Error bars represent 95 percent confidence intervals. Panel (B), example CE-optimal language (ambiguous) from a four-context simulation. Panel (C), example CE-optimal language (unambiguous) from a single-context simulation."}
img <- png::readPNG("figs/fig1.png")
grid::grid.raster(img)
```

@@

For comparison, we examine properties of optimal languages under two additional objectives. Zipf [-@Zipf1949a] proposed that the speaker-optimal language $\ell_{speaker}^*$ would minimize speaker effort and the listener-optimal language $\ell_{listener}^*$ would minimize listener effort. We define these objectives using the first and second half of equation 1 (see \href{https://bit.ly/2RBSGcU}{SI 2.2.}).\par

@@

Imagine a scenario in which a reader is beginning a news article. While they may have some knowledge about the article’s topic (perhaps from the title), they may not have complete knowledge of its contents, including the persons or events involved. In this setting, using a low-cost, but ambiguous referring expression (say a pronoun like "he") early may lead to misunderstanding if context is not informative. But, if by a later position enough contextual information has accumulated, it may be efficient to use the ambiguous expression. We pursue this general framework in Simulation 2 -- examining when in a discourse using ambiguity is efficient. We will consider “context” as analogous to a “topic” of conversation. \par

@@

We conduct $N=600$ simulations, generating discourses of length $|D|=30$ utterances, comparing three speaker models ($n=200$ each). We consider a single language $\ell$\footnote{See SI for the matrix notation of this language.} with $|U|=6$ and $|M|=4$ in which two of the utterances are ambiguous and lower cost than the unambiguous utterances. (Note that use of this particular language is not essential -- the results are broadly generalizable to languages that contain ambiguity, but exploring this space is computationally expensive.)\par

## Speaker agents
We vary the degree to which agents can use context for disambiguation. We consider three types of speaker models. Our \textit{Full pragmatics} agent, models a speaker who reasons about her listener and also has complete recall of the set of utterances in the discourse $D$. This speaker believes that the listener may not know the current topic $c_{current}$ at the start of the discourse, but can infer it over the discourse. We compare two baseline models. The first, a \textit{Partial pragmatics} baseline describes a speaker who reasons about a listener, but assumes they have no access to the discourse history. The second, a \textit{No pragmatics} baseline speaker does not consider a listener at all, but produces utterances according to the underlying language semantics ($\ell$) and topic probabilities ($p(M|C=c_{current}$) (see \href{https://bit.ly/2RBSGcU}{SI 3}).\par

@@

Figure 3, panel (A) shows the empirical probability a speaker uses an ambiguous utterance as a function of discourse position. The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse. The \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely. But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds. Panel (C) tracks our CE objective for each model over the discourse. Note that the objective decreases for all three models, primarily driven by the listeners updating their beliefs about the actual topic ($P(C=c_{current}|D)$). However, the objective declines more quickly under the \textit{Full} and \textit{Partial pragmatic} speakers as listener agents are better able to infer the correct context. Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs. Panel (B) tracks speaker effort, which remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines. But, effort declines in the \textit{Full pragmatic} model as speakers increasingly rely on ambiguous material later in the discourse.\par

@@

How do the competing pressures imposed by speakers and listeners give rise to the distributional regularities found in natural language? Zipf [-@Zipf1949a] proposed that the asymmetry between speaker and listener costs gives rise to a range of properties at the level of the lexicon. We explored the interactions of rational pragmatic agents as a framework for understanding efficient language structure and use. We focused on an argument on the communicative function of ambiguity [@Piantadosi2011a], deriving a novel speaker-listener Cross-Entropy objective for measuring the efficiency of linguistic systems from first principles of efficient language use. In Simulation 1 we showed that optimal languages are more likely to contain ambiguous material when context is informative. In Simulation 2 we showed how rational pragmatic agents use ambiguous material efficiently in conversation.\par

A limitation of the current work is an analysis of exactly how the CE objective compares to existing measures. For example, previous work has described competing speaker-listener pressures in terms of a trade-off of simplicity and informativeness [@KempRegier2012a] or expressivity and compressibility [@SmithTamarizKirby2013a] to explain linguistic regularities. Future work should assess the degree to which we can derive the same properties as previous studies using our current framework. More generally, we hope that this framework can serve as a domain general tool to assess the range of functionalist theories examining efficient language-structure and use.\par

\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {SI available at: \href{https://bit.ly/2RBSGcU}{\url{https://bit.ly/2RBSGcU}}. Code will be available upon acceptance.}}} \vspace{1em}
\noindent
